### 2023-04-03 paper-thesis/chapters/99-conclusion.tex
More feedback
Bit of work on the conclusion

This thesis presents an implementation of a dependently typed language in Statix. Our aim was to anther the following research questions:

@@

	\item Support for universe polymorphism~\cite[Universe Levels]{agda}, which allow functions and inductive data types to be polymorphic over a certain universe level. For example, this allows to define a \verb|List| data type that can be both a list of values and a list of types without defining it twice. We see no reason why this couldn't be cleanly implemented in Statix. 
	\item Another feature of Agda is implementing dependent pattern matching, as described in "Eliminating Pattern Matching without K"~\cite{patternmatching}. This is requires a desugaring stage, in which matches are desugared to eliminators, which should be implemented in Stratego before type checking. This should also be possible to implement, but it may be quite a bit of work.
	\item Inference in our language is currently implemented by explicitly specifying a placeholder \verb|_| where inference is required. In Agda we can specify an implicit argument, which is automatically inferred unless explicitly provided. We see no reason why this couldn't be cleanly implemented in Statix. 

@@

	\item Agda has holes, which allows the user of the language to build proofs interactively. One of these interactions is proof search, when triggered Agda will try to find a term that fits the expected type of a hole. The interactive part of this was discussed in chapter~\ref{ch:relatedwork}, this is not yet supported by Spoofax. After support for this is added, a proof search algorithm could be implemented in Stratego, such as the structurally recursive unification algorithm described by McBride~\cite{proofsearch}.
	\item Agda allows function arguments to be marked as irrelevant, meaning that the value will be erased at runtime. An even stronger version of this feature is implementing a quantitive type system, such as in Idris~2~\cite{idris2}, which combines linear and dependent types. Implementing this in Spoofax should be possible, but it may be quite a bit of work as it is a complex type system.

@@

\begin{itemize}
	\item There has not yet been an implementation of a sub-structural type system in Spoofax. A sub-structural type system is a type system that places restrictions on how often variables can be used. Now that languages with sub-structural type systems, such as Rust, are becoming popular, we should see if the goal of Statix to cover a broad range of type systems, still holds true for these. 
	\item Effect systems are a formal system that describe the side effects of a program. These allow programmers to specify what side effects a function has, which can then be verified by a type checker. Researching whether an effect system could be cleanly specified in Statix may lead to more insights on the limits of Statix.
\end{itemize}

>>>>

This thesis presents an implementation of a dependently typed language in Statix. Our aim was to answer the following research questions:

@@

	\item Support for universe polymorphism~\cite[Universe Levels]{agda}, which allow functions and inductive data types to be polymorphic over a certain universe level. For example, this allows to define a \verb|List| data type that can be both a list of values and a list of types without defining it twice. Agda has quite a rather involved definition of equality between universe levels, which may be difficult to implement in Statix, but at least a simple version of universe polymorphism should be possible to implement in Statix.
	\item Another feature of Agda is implementing dependent pattern matching, as described by Coquand~\cite{patternmatching}. Dependent pattern matching is complex because it requires interleaved type checking and desugaring, which is problematic since desugaring is usually done in Stratego, while type checking is done in Statix. The solution may be to do the desugaring in Statix instead, but at some point this is not ideal, since Statix is not meant for this. More research is needed on a good solution for this problem.
	\item Inference in our language is currently implemented by explicitly specifying a placeholder \verb|_| where inference is required. In Agda we can specify an implicit argument, which is automatically inferred unless explicitly provided. It can be tricky to determine when an implicit argument has to be inserted, since this may require some type information~\cite{implicit_arguments}. More research is needed on whether the solution from this paper can be implemented in Statix.

@@

	\item Agda has holes, which allows the user of the language to build proofs interactively. One of these interactions is proof search, when triggered Agda will try to find a term that fits the expected type of a hole. The interactive part of this was discussed in chapter~\ref{ch:relatedwork}, this is not yet supported by Spoofax. After support for this is added, a proof search algorithm could be implemented in Stratego, such as the structurally recursive unification algorithm described by McBride~\cite{proofsearch}. Running the algorithm may require running type checking code, so here we encounter the interleaving problem again. More research is needed on whether proof search can be implemented in Spoofax.
	\item \todo{irrelevance}
	\item Agda allows function arguments to be marked as erased, meaning that the value will be erased at runtime. An even stronger version of this feature is implementing a quantitive type system, such as in Idris~2~\cite{idris2}, which combines linear and dependent types. Implementing this in Spoofax should be possible, but it may be quite a bit of work as it is a complex type system.

@@

\begin{itemize}
	\item There has not yet been an implementation of a sub-structural type system in Spoofax. A sub-structural type system is a type system that places restrictions on how often variables can be used. Now that languages with sub-structural type systems, such as Rust, are becoming popular, we should see if the goal of Statix to cover a broad range of type systems, still holds true for these. 
	\item Effect systems are a formal system that describe the side effects of a program. These allow programmers to specify what side effects a function has, which can then be verified by a type checker. Researching whether an effect system could be cleanly specified in Statix may lead to more insights on the limits of Statix.
	\item \todo{quantitive type system}
\end{itemize}
